\chapter{Conclusions and future work}

We have successfully developed a competitive chess engine based on an enhanced minimax search with alpha-beta pruning, a straightforward materialistic evaluation, and a Most Valuable Victim-Least Valuable Aggressor (MVV-LVA) move-ordering heuristic.

\vspace{1em}

\noindent Building on this basic implementation, we have researched and analyzed the following algorithmic techniques to further improve the bot's playing strength:

\begin{itemize}[itemsep=1pt]

    \item \textit{Transposition tables with Zobrist hashing}: provided an increase in performance by avoiding redundant position evaluations.
    
    \item \textit{Move generator with magic bitboards and PEXT instructions}: a lookup table was used to retrieve legal moves for pieces with constant-time access. The table index is computed using the hardware-accelerated PEXT instruction when available; otherwise, the engine falls back to the magic bitboards technique.
    
    \item \textit{Evaluation with king safety and piece mobility parameters}: introduced additional parameters to evaluate the position, incorporating king safety and piece mobility. This showed no clear improvement, likely due to the added computational cost and the difficulty of directly correlating these factors with positional strength.
    
    \item \textit{Multithreaded search}: requires further optimization of data structures and algorithms for concurrent access to yield tangible gains.
    
    \item \textit{Search with Late Move Reductions}: did not improve performance, as our current move-ordering heuristic is not strong enough to support such aggressive pruning.
\end{itemize}

\noindent The engine achieved an Elo rating of 1900 on \textit{Lichess} while running on a Raspberry Pi 5 with a 2GB transposition table, demonstrating its efficiency even on resource-constrained hardware.

\vspace{1em}

\noindent The engine's \textit{Lichess} profile can be found at:\\
\url{https://lichess.org/@/AlphaDeepChess}


\newpage
\section{Future work}

\noindent Potential avenues for further improvement include:

\begin{itemize}[itemsep=1pt]
    \item \textit{Neural-network evaluation and move ordering}:  
    Application of neural networks for the evaluation function and the move ordering heuristic, following the steps of \textit{Stockfish}, the actual best chess engine. This could unlock more aggressive pruning strategies.
    \item \textit{Advanced pruning techniques}:  
    With a stronger evaluation and ordering heuristic, revisit and tune the actual Late Move Reductions, and implement null move pruning technique.
    \item \textit{Multithreading optimizations}:  
    Refine concurrent search algorithm and redesign shared data structures like transposition tables to support concurrent access.
\end{itemize}
